Programming Language Foundations in Agda

### 2019-07-14 papers/scp/PLFA.tex
new material on animation

@@

It is immediately obvious that progress and preservation make it
trivial to construct a prototype evaluator, and yet we cannot find such
an observation in the literature nor mentioned in an introductory
text.  It does not appear in SF, nor in \citet{Harper-2016}.  A plea
to the Agda mailing list failed to turn up any prior mentions.
The closest related observation I have seen in the published
literature is that evaluators can be extracted from proofs of
normalisation \citep{Berger-1993,Dagand-and-Scherer-2015}.

(Late addition: Our plea to the Agda list eventually bore fruit.  Oleg
Kiselyov directed me to unpublished remarks on his web page where he
uses the name \texttt{eval} for a proof of progress and notes ``the
very proof of type soundness can be used to evaluate sample
expressions'' \citep{Kiselyov-2009}.)

>>>>

% new
Philip had been exposed to the work of the K team, as both consulted
for IOHK, a cryptocurrency firm.  This put us keenly in mind of the
need for animation; Philip sometime referred to this as ``K-envy'' or
``Redex-envy''.

@@

It is immediately obvious that progress and preservation make it
trivial to construct a prototype evaluator, and yet we cannot find such
an observation in the literature nor mentioned in an introductory
text.  It does not appear in SF, which introduces a specialised
\texttt{normalise} tactic instead.  A plea
to the Agda mailing list failed to turn up any prior mentions.
The closest related observation we have seen in the published
literature is that evaluators can be extracted from proofs of
normalisation \citep{Berger-1993,Dagand-and-Scherer-2015}.

% NEW
Some researchers are clearly familiar with the connection between
progress and preservation and animation.  In private correspondence,
Bob Harper referred to it as the \emph{pas de deux}, a dance between
progress, which takes well-typing to a step, and preservation, which
takes a step back to well-typing.  Nonetheless, neither the technique
nor the appealing terminology, appears in \citet{Harper-2016}.  The
appeal to the Agda mailing list bore late fruit: Oleg Kiselyov
directed me to unpublished remarks on his web page where he uses the
name \texttt{eval} for a proof of progress and notes ``the very proof
of type soundness can be used to evaluate sample expressions''
\citep{Kiselyov-2009}.  Nonetheless, as of this writing, we still have
not located a mention in the published literature.

% NEW
There are places in the literature where one might expect a remark on
the relation between progress and preservation and animation---but no
such remark appears.  In the PoplMark Challenge
\citep{Ayedemir-et-al-2005}, Challenge~2A is to prove progress and
preservation for System F$_{<:}$, while Challenge~3 is to prove
animation for the same system. Nowhere do the authors indicate that in
an intuitionistic logic these are essentially the same problem.
\cite{Owens-et-al-2016}, when discussing extraction of animators for
small-step semantics, mention Redex and K, but no other possibilities.
We hope the stress in PLFA on the fact that in an intuitionistic
setting progress and preservation imply animation will mean that the
connection becomes more widely known.
